---
title: "PSYR6003-Assignment1-NBasso"
author: "Nicole Basso"
format: docx
editor: visual
---

## Load Libraries Needed For Assignment

```{r}
library(tidyverse)
library(janitor)
library(stats)
library(skimr)
library(car)
```

## Question 1

## Load "Avengers" CSV file dataset, assigning it to object "avengers"

## then previewing the dataset using "glimpse" for inspection of variables

## Glimpse showed there was no Participant ID column included in the original dataset; adding an ID column "PID" at the beginning of the dataset to aid in Data Inspection & Cleaning

```{r}
avengers <- read.csv("avengers.csv") %>% 
            glimpse() %>% 
            mutate(PID = row_number(), .before = 1)
```

## Checking for unusual observations in the dataset for all columns except "PID": (1) Missing/Empty/String typos Values \[identified by "skim_tee"\]; (2) Outliers/Illogical negative or zero values \[IQR summary with "skim_tee"\]; (3) Duplicate Observations \[identified by "get_dupes"\]

```{r}
avengers %>% 
  skim_tee(-PID) %>% 
  get_dupes(-PID)
```

## Cleaning Character Variables

## The three character variables should all have three unique strings and no empty values

## superpower and died should only have options yes, no, or NA

## north_south should have only options north, south, or NA

## Assessing current strings for typos/empty cells are coercing them into labelled factors with only the above options under new object "clean.avengers"

## labels will be the denoted options wherein "yes" and "north" correspond to "1", "no" and "south" correspond to "0", and missing values correspond to "NA"

## Checking recode by looking at unique values for each as the first 6 rows of the dataset

```{r}
print(unique(avengers$superpower))
print(unique(avengers$north_south))
print(unique(avengers$died))

clean.avengers <- mutate(avengers, 
            superpower = Recode(avengers$superpower, 
              "c('yes', '\"yes\"') = 'yes'; 'no' = 'no'; else = NA", as.factor = T, 
              levels = c("yes", "no", NA)),
            north_south = Recode(avengers$north_south, 
                "'north' = 'north'; 'south' = 'south'; else = NA",  as.factor = T,
                levels = c("north", "south", NA)),
            died = Recode(avengers$died, 
                "'yes'  =  'yes'; 'no' = 'no'; else = NA", as.factor = T, 
                levels = c("yes", "no", NA)),
            .keep = "all")

print(unique(clean.avengers$superpower))
print(unique(clean.avengers$north_south))
print(unique(clean.avengers$died))

head(clean.avengers)

```


## Subsetting rows with unusual values identified in the previous chunk into a new tibble called "sus" in order to manually inspect flagged rows

## those with empty/missing values, extreme values, and/or negative values will be subsetted; no duplicate rows were found and thus code for duplication subsetting will not be written

## extreme values are identified as 1.5xIQR in accordance with Tukey's Method

```{r}

sus <- clean.avengers %>% 
            filter(if_any(where(is.numeric), ~ . <= 0))

sus <- avengers %>% 
          filter(!complete.cases(avengers), if_any(where(is.numeric), ~ . <= 0), 
          if_any(where(is.numeric), ~ . <= 1.5*IQR()), if_any(where(is.numeric), ~ . >= 1.5*IQR(~ .)))

```

## Question 2

## Subsetting the Data to Include only Complete clean cases


